---
title: "HW5"
output: pdf_document
date: "2023-04-01"
---

```{r}
# Import data set 
train <- read.delim("vowel.train.txt", header = TRUE, sep = ",")
train = train[-1]
str(train)
train$y = as.factor(train$y)
test <- read.delim("vowel.test.txt", header = TRUE, sep = ",")
test = test[-1]
test$y = as.factor(test$y)
```

Fit a random forest or gradient boosted model to the “vowel.train” data using all of the 11 features using the default values of the tuning parameters.

```{r, message=FALSE, echo=FALSE}
library(randomForest)

# Fit a random forest model with default parameters
fit_rf <- randomForest(y ~ ., data = train)
```

Use 5-fold CV to tune the number of variables randomly sampled as candidates at each split if using random forest, or the ensemble size if using gradient boosting.

```{r,message=FALSE, echo=FALSE, warning=FALSE}
library(caret)
# Set up the training control for cross-validation
train_control <- trainControl(method = "cv", number = 5)

# Tune the number of variables randomly sampled as candidates at each split
tune_grid <- expand.grid(mtry = 1:11)
set.seed(123)
fit_rf_tuned <- train(y ~ ., data = train, method = "rf",
                      importance = TRUE, proximity=TRUE,
                      tuneGrid = tune_grid, trControl = train_control)

fit_rf_tuned$bestTune # 1 variable
(fit_rf_tuned$finalModel)

print(fit_rf_tuned)          ## summary of fit object
plot(fit_rf_tuned)           ## plot OOB MSE as function of # of trees
```

With the tuned model, make predictions using the majority vote method, and compute the misclassification rate using the ‘vowel.test’ data.
```{r, message=FALSE}
# Make predictions on the test data using the tuned model
pred_rf <- predict(fit_rf_tuned, newdata = test)

# Compute the misclassification rate
(misclassification_rate <- mean(pred_rf != test$y))
```
